Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2016-09-29 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
working on conclusions
skipped ahead

\clearpage
Abbreviations
\begin{tabbing}
\hspace{1.5in} \= \\
EMS \> Emergency Medical Services \\
FDS \> Fire Dynamics Simulator \\
HGL \> Hot Gas Layer \\
HRR \> Heat Release Rate \\
HRRPUA \> Heat Release Rate per Unit Area \\
NIST \> National Institute of Standards and Technology \\
\end{tabbing}

@@

@@

@@

@@

Repeatibilty
Energy to Damage
Flame Height
Modeling

>>>>

\clearpage
Abbreviations
\begin{tabbing}
\hspace{1.5in} \= \\
EMS \> Emergency Medical Services \\
FDS \> Fire Dynamics Simulator \\
HGL \> Hot Gas Layer \\
HRR \> Heat Release Rate \\
HRRPUA \> Heat Release Rate per Unit Area \\
NFPA \> National Fire Protection Association  \\
NIST \> National Institute of Standards and Technology \\
\end{tabbing}

@@

distance from the wall

@@

duration

@@

corners

@@

The main objective of the research was to examine the repeatability of pre-flashover fire patterns generated from exposure to short duration (300 s maximum) well characterized fires.  The first step was to exmaine the repeatability of different pre-flashover fires.  Three different fuels were used: natural gas, gasoline, and polyurethane foam.  Each fuel had a similar top surface area.  The heat release rate data showed that the variability was greater for more complex fuels.  While the variation in peak heat release rate with the natural gas was similar to the expanded uncertainty of the measurement system, $\pm$ 11~\%, the variation in the peak heat release rates of the gasoline and the polyurethane foam increased due to increased uncertainties in the burning behavior of the fuels.  The fire pattern heights generated on painted gypsum wallboard by the natural gas burner and gasoline fueled fires were shown to have uncertainties of $\pm$ 18~\% or less based on a Type A statistical analysis with 95~\% confidence limits. The patterns generated by the polyurethane foam fueled fire, had more variability than the natural gas and the gasoline fire patterns.  This is likely due to the greater variability of the burning of the solid fuel and the lower heat release rate.    The comparison of the fire pattern heights and the mean flame height demonstrated that the “steady state” natural gas fires exhibited the highest level of agreement and polyurethane foam fueled fire exhibited the least agreement, with the gasoline fueled fires in between.  This trend is the same as exhibited by the repeatability of the heat release rate measurements.  Given that heat release rate is the source of the energy which is transferred to the walls in order generate the fire patterns, it stands to reason that the greater the uncertainty in the heat release rate the greater the unceratinty in pattern and any associated thermal measurements.

This paper also highlighted the many uncertainties that need to be considered in what appear to be simple fire experiments. There were a variety of uncertainties related to the composition of the fuels, measurement methods, and analysis techniques just to list a few of the areas considered.

The second objective of this study was to examine the capability of numerical models to simulate the fires and the resulting fire patterns based on input data collected from engineering reference sources, and bench-scale and full-scale fire experiments conducted in this study.  The numerical tools used ranged from simple empirical algorithms to a computational fluid dynamics model.  Each of the tools are freely available to investigators and many of the tools have been packaged in a form that allows the user to input the ``data'' and generate a result.  There is software available to aid with the development of the computational fluid dynamics model.  The first challenge in using the numerical tools is finding the appropriate data to us as input.
